\section{Contributions}
\label{sec:contrib}

In this section we discuss the unique contributions of the underlying work \cite{sanchez2019can}.

\paragraph{Tracking.}

\citeauthor{sanchez2019can} extend the contributions of previous work about the GDPR and cookie control. They show that
the GDPR has a global impact from the perspective of a user from the EU whereas related work only looks at the impact on
websites in the EU from the perspective of users from within and outside the EU. Another interesting novelty is the
detection method for uniquely identifiable tracker cookies. Typically, trackers are detected by known tracker lists.
Other papers do not differentiate between tracker cookies and non-tracker cookies, but consider the general number of
cookies. \citeauthor{sanchez2019can} detect tracker cookies by considering whether a cookie could be used to uniquely
identify a user via the entropy of the value of the cookie. With this technique, they can identify trackers which are not on known
tracker lists and first party trackers which otherwise could be overseen.

\paragraph{Privacy Policies.}

\citeauthor{sanchez2019can} evaluate privacy policies of the top websites with the \emph{Flesch Reading Ease Score} (FRES)~\cite{flesch1948new}
and the \emph{Flesch-Kincaid Reading Level} (FKRL)~\cite{kincaid1975derivation}.
This was done before the GDPR and \citeauthor{sanchez2019can} compare the previous results to the results after the
GDPR. They show that privacy policies have gotten more readable after the GDPR and that more websites provide a privacy
policy at all% .

